### 2024-02-05 docs/projects/noj/ggplot.qmd
Noj - rerendered

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="ggplot_files/md-default5.js" type="text/javascript"></script><script src="ggplot_files/md-default6.js" type="text/javascript"></script>

@@

# Exploring ggplot
following
[cxplot](https://cxplot.com/index.html)'s internal ggplot.as.list
in representing a plot as a data structure.

@@

@@

::: {.sourceClojure}
```clojure
(def plot
  (r "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))"))
```
:::

::: {.sourceClojure}
```clojure
(-> plot
    plotting/plot->svg
    kind/html)
```
:::

@@

<g id="surface428">

@@

::: {.sourceClojure}
```clojure
(-> plot
    r->clj
    (dissoc :data))
```
:::

::: {.printedClojure}
```clojure
{:layers [nil nil],
 :scales nil,
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates nil,
 :facet nil,
 :plot_env nil,
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

::: {.sourceClojure}
```clojure
(defn gg-facet [ggplot]
  (let [f (-> ggplot
              (r$ 'facet)
              (r$ 'params)
              (r$ 'facets))]
    (when (-> `(is.null ~f)
              r
              r->clj)
      (let [facet (r `(ls ~f))]
        {:facet facet
         :facet-levels (-> plot
                           (r$ 'data)
                           (r/bra facet))}))))
```
:::

::: {.sourceClojure}
```clojure
(-> plot
    gg-facet
    (update-vals r->clj))
```
:::

```{=html}
<div><p>{</p><div style="margin-left:10%;width:110%;"><div><pre><code class="sourceCode language-clojure printed-clojure">:facet ["cyl"]</code></pre></div><table><tr><td valign="top"><div><pre><code class="sourceCode language-clojure printed-clojure">:facet-levels
</code></pre></div></td><td><div style="margin-left:10px;"><span data-qmd="_unnamed [234 1]:

| :cyl |
|-----:|
|    4 |
|    4 |
|    4 |
|    4 |
|    6 |
|    6 |
|    6 |
|    4 |
|    4 |
|    4 |
|  ... |
|    4 |
|    4 |
|    5 |
|    5 |
|    4 |
|    4 |
|    4 |
|    4 |
|    6 |
|    6 |
|    6 |

"></span></div></td></tr></table></div><p>}</p></div>
```

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"
    r
    (ggolot->clj {:avoid #{"data" "plot_env"}}))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

@@

::: {.sourceClojure}
```clojure
(defn ggplot-summary [r-code]
  (let [plot (r r-code)
        clj (-> plot
                (ggolot->clj {:avoid #{"data" "plot_env"}})
                (->> (walk/postwalk (fn [form]
                                      (if (and (symbol? form)
                                               (-> form str (= "~")))
                                        (str form)
                                        form)))))]
    (kind/fragment
     [(h4 "code")
      (kind/md
       (format "\n```{r eval=FALSE}\n%s\n```\n"
               r-code))
      (h4 "image")
      (plotting/plot->buffered-image plot)
      (h4 "clj")
      (kind/pprint clj)])))
```
:::

```{=html}
<h3>code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy))
         + geom_point())
```

```{=html}
<h3>image</h3>
```

```{=html}
<h3>clj</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x ["~" cty], :y ["~" hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

## Exlploring a few plots

@@

```{=html}
<h3>code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy))
         + geom_point())
```

```{=html}
<h3>image</h3>
```

```{=html}
<h3>clj</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x ["~" cty], :y ["~" hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

@@

```{=html}
<h3>code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point())
```

```{=html}
<h3>image</h3>
```

```{=html}
<h3>clj</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x ["~" cty], :y ["~" hwy], :colour ["~" [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

```{=html}
<h3>code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method="lm"))
```

```{=html}
<h3>image</h3>
```

```{=html}
<h3>clj</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method}}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x ["~" cty], :y ["~" hwy], :colour ["~" [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

```{=html}
<h3>code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method="lm")
         + facet_wrap(~cyl))
```

```{=html}
<h3>image</h3>
```

```{=html}
<img src="ggplot_files/4.png" style="width:100%;" />
```

```{=html}
<h3>clj</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method}}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x ["~" cty], :y ["~" hwy], :colour ["~" [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl ["~" cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels &lt;- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n&lt;bytecode: 0x55eafa698ac0&gt;\n&lt;environment: namespace:ggplot2&gt;\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

>>>>

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="ggplot_files/md-default4.js" type="text/javascript"></script><script src="ggplot_files/md-default5.js" type="text/javascript"></script>

@@

# Exploring ggplot - DRAFT
Here we explore [ggplot2](https://ggplot2.tidyverse.org/) from Clojure. The goal is to eventually implement a similar grammar in Clojure.

## Setup
We will use [ClojisR](https://scicloj.github.io/clojisr/) to call R from Clojure.

@@

Loading ggplot2:

@@

The base R library:

## An example plot
Here is how we may generate and display a plot:

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))"
    r
    plotting/plot->svg
    kind/html)
```
:::

@@

<g id="surface679">

@@

We may also [generate](https://scicloj.github.io/clojisr/doc/clojisr/v1/codegen-test/) the R code from a Clojure form and pass Clojure data to R, but for now, we are simply using R code using an R dataset.
We are using the `mpg` dataset, which is part of the `ggplot` package.

::: {.sourceClojure}
```clojure
(-> "mpg"
    r
    r->clj)
```
:::

_unnamed [234 11]:

| :manufacturer |     :model | :displ | :year | :cyl |     :trans | :drv | :cty | :hwy | :fl |     :class |
|---------------|------------|-------:|------:|-----:|------------|------|-----:|-----:|-----|------------|
|          audi |         a4 |    1.8 |  1999 |    4 |   auto(l5) |    f |   18 |   29 |   p |    compact |
|          audi |         a4 |    1.8 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   p |    compact |
|          audi |         a4 |    2.0 |  2008 |    4 | manual(m6) |    f |   20 |   31 |   p |    compact |
|          audi |         a4 |    2.0 |  2008 |    4 |   auto(av) |    f |   21 |   30 |   p |    compact |
|          audi |         a4 |    2.8 |  1999 |    6 |   auto(l5) |    f |   16 |   26 |   p |    compact |
|          audi |         a4 |    2.8 |  1999 |    6 | manual(m5) |    f |   18 |   26 |   p |    compact |
|          audi |         a4 |    3.1 |  2008 |    6 |   auto(av) |    f |   18 |   27 |   p |    compact |
|          audi | a4 quattro |    1.8 |  1999 |    4 | manual(m5) |    4 |   18 |   26 |   p |    compact |
|          audi | a4 quattro |    1.8 |  1999 |    4 |   auto(l5) |    4 |   16 |   25 |   p |    compact |
|          audi | a4 quattro |    2.0 |  2008 |    4 | manual(m6) |    4 |   20 |   28 |   p |    compact |
|           ... |        ... |    ... |   ... |  ... |        ... |  ... |  ... |  ... | ... |        ... |
|    volkswagen | new beetle |    2.0 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   r | subcompact |
|    volkswagen | new beetle |    2.0 |  1999 |    4 |   auto(l4) |    f |   19 |   26 |   r | subcompact |
|    volkswagen | new beetle |    2.5 |  2008 |    5 | manual(m5) |    f |   20 |   28 |   r | subcompact |
|    volkswagen | new beetle |    2.5 |  2008 |    5 |   auto(s6) |    f |   20 |   29 |   r | subcompact |
|    volkswagen |     passat |    1.8 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   p |    midsize |
|    volkswagen |     passat |    1.8 |  1999 |    4 |   auto(l5) |    f |   18 |   29 |   p |    midsize |
|    volkswagen |     passat |    2.0 |  2008 |    4 |   auto(s6) |    f |   19 |   28 |   p |    midsize |
|    volkswagen |     passat |    2.0 |  2008 |    4 | manual(m6) |    f |   21 |   29 |   p |    midsize |
|    volkswagen |     passat |    2.8 |  1999 |    6 |   auto(l5) |    f |   16 |   26 |   p |    midsize |
|    volkswagen |     passat |    2.8 |  1999 |    6 | manual(m5) |    f |   18 |   26 |   p |    midsize |
|    volkswagen |     passat |    3.6 |  2008 |    6 |   auto(s6) |    f |   17 |   26 |   p |    midsize |

## Representing plots as Clojure data
Inspired by
[cxplot](https://cxplot.com/index.html)'s internal ggplot.as.list finction, let us represent ggplot objects as R data structures.
A ggplot object is an R list of [ggproto](https://bookdown.dongzhuoer.com/hadley/ggplot2-book/introducing-ggproto) objects. We recursively unwrap this structure and convert it to Clojure.

In the conversion, we avoid some big parts of the structure, e.g., the `"plot_env"` member. We also do no report the toplevel `"data"` member, which is simply the dataset.
For example:

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"
    r
    (ggolot->clj {:avoid #{"plot_env"}})
    (dissoc :data))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

@@

## Exlploring a few plots
Let us explore and compare a few plots this way:

@@

::: {.sourceClojure}
```clojure
(defn ggplot-summary [r-code]
  (let [plot (r r-code)
        clj (-> plot
                (ggolot->clj {:avoid #{"plot_env"}})
                (dissoc :data))]
    (kind/fragment
     [(h4 "R code")
      (kind/md
       (format "\n```\n%s\n```\n"
               r-code))
      (h4 "plot")
      (-> plot
          plotting/plot->buffered-image)
      (h4 "clj data")
      (kind/pprint clj)])))
```
:::

@@

```{=html}
<h3>R code</h3>
```

```
(ggplot(mpg, aes(cty, hwy))
         + geom_point())
```

```{=html}
<h3>plot</h3>
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

@@

```{=html}
<h3>R code</h3>
```

```
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point())
```

```{=html}
<h3>plot</h3>
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

```{=html}
<h3>R code</h3>
```

```
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method="lm"))
```

```{=html}
<h3>plot</h3>
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

```{=html}
<h3>R code</h3>
```

```
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method="lm")
         + facet_wrap(~cyl))
```

```{=html}
<h3>plot</h3>
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~ cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels &lt;- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n&lt;bytecode: 0x559b84332b20&gt;\n&lt;environment: namespace:ggplot2&gt;\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

### 2024-02-05 docs/projects/noj/ml.qmd
Noj - rerendered

Residuals:
       Min          1Q      Median          3Q         Max
   -0.8799     -0.2284      0.0138      0.2569      0.6354

Coefficients:
                  Estimate Std. Error    t value   Pr(>|t|)
Intercept           1.9687     0.3043     6.4705     0.0000 ***
sepal-width         0.6222     0.0800     7.7748     0.0000 ***
petal-length        0.6929     0.0713     9.7167     0.0000 ***
petal-width        -0.5072     0.1639    -3.0944     0.0026 **
---------------------------------------------------------------------
Significance codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 0.3196 on 96 degrees of freedom
Multiple R-squared: 0.8577,    Adjusted R-squared: 0.8533
F-statistic: 192.9066 on 4 and 96 DF,  p-value: 1.640e-40

>>>>

Residuals:
       Min          1Q      Median          3Q         Max
   -0.7420     -0.2582      0.0213      0.2341      0.8930

Coefficients:
                  Estimate Std. Error    t value   Pr(>|t|)
Intercept           1.8651     0.3225     5.7825     0.0000 ***
sepal-width         0.6599     0.0863     7.6500     0.0000 ***
petal-length        0.7003     0.0701     9.9877     0.0000 ***
petal-width        -0.5694     0.1585    -3.5924     0.0005 ***
---------------------------------------------------------------------
Significance codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 0.3268 on 96 degrees of freedom
Multiple R-squared: 0.8465,    Adjusted R-squared: 0.8418
F-statistic: 176.5359 on 4 and 96 DF,  p-value: 6.132e-39
